Ultrafast manipulation of Heisenberg exchange and Dzyaloshinskii–Moriya interactions in antiferromagnetic insulators

### 2019-01-29 Paper/paper.tex
wrap on

Where $t_{ij}^{\sigma}$ are the corresponding amplitudes for the unperturbed Hamiltonian, i.e. $t_{ij}^{\sigma} = t_1$ for $i,j$ being NN and $t_{ij}^{\sigma} = t_2 + i\Delta\nu_{ij}\sigma^z_{\sigma, \sigma}$ for $i,j$ being NNN. We can obtain the spin Hamiltonian by using the relations:
\begin{align}
\hat{c}_{i \sigma}^\dagger \hat{c}_{i \sigma'} &= \delta_{\sigma \sigma'} \frac{1}{2} (n_{i \uparrow} + n_{i \downarrow}) + \boldsymbol{S}_i\boldsymbol{\sigma}_{\sigma', \sigma} \label{SpinOperatorInv1}\\ 
\hat{c}_{i \sigma} \hat{c}_{i \sigma'}^\dagger &= \delta_{\sigma \sigma'} \frac{1}{2} (2 - n_{i \uparrow} - n_{i \downarrow}) - \boldsymbol{S}_i\boldsymbol{\sigma}_{\sigma, \sigma'} \label{SpinOperatorInv2}
\end{align}
and $n_{i \uparrow} + n_{i \downarrow} = 1$.

\begin{subsection}{Kane--Mele--Hubbard model}

The first model to describe topological insulators was introduced by Kane and Mele \cite{Kane2005} to describe quantum spin Hall effect in graphene. In a honeycomb lattice time reversal symmetry and inversion symmetry allow only next-nearest neighbor spin orbit coupling, which is known as intrinsic spin orbit coupling. In these circumstances the system can be modeled by the Kane-Mele-Hubbard model, which has hopping amplitudes:

\begin{equation}
t_{ij}^{\sigma \sigma'} = \begin{cases}
	\delta_{\sigma \sigma'}t_1 & \text{for } i, j \text{ nearest neighbors} \\
	-\Delta \nu_{ij} \hat{e}_z \cdot \bs{\sigma}_{\sigma, \sigma'} & \text{for } i, j \text{ next nearest neighbors} \\
	0 & \text{ otherwise} \n
\end{cases} \quad
\end{equation}

Where $\Delta$ is the intrinsic spin orbit coupling constant. $\nu_{ij}=\pm 1$ depending on whether the electron traversing from $i$ to $j$ makes a right ($+1$) or a left turn ($-1$). Then, introducing the spin operators and the corresponding hopping amplitudes in (\ref{GeneralHeff}) we obtain the effective spin Hamiltonian of this model:

\begin{equation}
\label{KMHeff}
\hat{H}_{\text{KMH}}^{\text{eff}} = \sum_{\langle i,j \rangle} J_{ij} \bs{S}_i \bs{S}_j + \sum_{\langle \langle i,j \rangle \rangle} \bs{S}_i \bs{\Gamma}_{ij} \bs{S}_j 
\end{equation}

Where the exchange interaction couples NN spins and an anisotropic exchange interaction couples NNN spins:

\begin{align*}
J_{ij} &= 2t_1^2\frac{\mathcal{J}_{n}^2(\alpha_{ij})}{\text{U}+n\omega} \\
\bs{\Gamma}_{ij} &= 2\Delta^2 \text{diag}(-1,-1,1)\frac{\mathcal{J}_{n}^2(\alpha_{ij})}{\text{U}+n\omega}
\end{align*}
\end{subsection}

\begin{subsection}{Modified Kane--Mele--Hubbard model}

As we saw in the previous section the effective Hamiltonian \ref{KMHeff} does not contain NNN DMI terms. The reason for this is because DMI originates from the virtual hoppings $i\rightarrow j\rightarrow i$ combining SOI hopping and kinetik hopping, since in the original Kane-Mele model we don't find NNN kinetik hopping, the DMI term vanishes. Next we will study the same Hamiltonian adding a finite NNN hopping term $t_2$, which can be understood as a second order NN hopping process (the direct NNN hopping integral would usually be much smaller). Thus we start with:

\begin{align}
\label{MKMH}
\hat{H} &= - t_1\sum_{\langle i,j \rangle, \sigma} \hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} + \n \\
	&+\sum_{\langle \langle i,j \rangle \rangle, \sigma}(t_2 + i\Delta\nu_{ij}\sigma^z_{\sigma, \sigma})\hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma} + 
	\text{U}\hat{D}
\end{align}

Proceeding in the same way we find the effective spin Hamiltonian:

\begin{align}
\label{MKMHeff}
&\hat{H}_{\text{eff}}(t) = \sum_{\langle i,j \rangle} J_{1,ij}\bs{S}_i\bs{S}_j +\n \\
&+ \sum_{\langle \langle i,j \rangle \rangle} \left\{ J_{2,ij}\bs{S}_i\bs{S}_j + \bs{D}_{2,ij} \bs{S}_i \times \bs{S}_j + \bs{S}_i \bs{\Gamma}_{ij} \bs{S}_j \right\}
\end{align}

@@

\end{subsection}

>>>>

Where $t_{ij}^{\sigma}$ are the corresponding amplitudes for the unperturbed Hamiltonian, i.e. $t_{ij}^{\sigma} = t_1$ for $i,j$ being NN and $t_{ij}^{\sigma} = t_2 + i\Delta\nu_{ij}\sigma^z_{\sigma, \sigma}$ for $i,j$ being NNN. We can obtain the spin Hamiltonian by using the relations:
\begin{align}
\hat{c}_{i \sigma}^\dagger \hat{c}_{i \sigma'} &= \delta_{\sigma \sigma'} \frac{1}{2} (n_{i \uparrow} + n_{i \downarrow}) + \boldsymbol{S}_i\boldsymbol{\sigma}_{\sigma', \sigma} \label{SpinOperatorInv1}\\ 
\hat{c}_{i \sigma} \hat{c}_{i \sigma'}^\dagger &= \delta_{\sigma \sigma'} \frac{1}{2} (2 - n_{i \uparrow} - n_{i \downarrow}) - \boldsymbol{S}_i\boldsymbol{\sigma}_{\sigma, \sigma'} \label{SpinOperatorInv2}
\end{align}
and $n_{i \uparrow} + n_{i \downarrow} = 1$. 
and summing over the spin states. The resulting effective spin Hamiltonian is:

\begin{align}
\label{MKMHeff}
&\hat{H}_{\text{eff}} = \sum_{\langle i,j \rangle} J_{1,ij}\bs{S}_i\bs{S}_j +\n \\
&+ \sum_{\langle \langle i,j \rangle \rangle} \left\{ J_{2,ij}\bs{S}_i\bs{S}_j + \bs{D}_{2,ij} \bs{S}_i \times \bs{S}_j + \bs{S}_i \bs{\Gamma}_{ij} \bs{S}_j \right\}
\end{align}

@@
